### 2022-12-29 paper.Rmd
minor

When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups, i.e., a which partition where both the standard $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ are close to the global optimum value. However, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree, making a tradeoff necessary [@ferligoj1992direct]. A motivating example illustrates that in the case of anticlustering, aiming to simultaneously optimize similarity of means and variances may however be feasible. I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups. Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions. To divide the 14 elements into two groups that are similar with regard to their mean and variance, a partition should be chosen that is depicted in the upper right corner of the scatterplot. The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$. Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize $\mathit{SSE}$ at the same time. The diamond symbol represents the partition that maximizes $\mathit{SSE_{kplus}}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$. This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$. Thus, the optimal partition according to $\mathit{SSE_{kplus}}$ fulfills the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ to an astonishing degree.

```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups. The partition represented by the triangular minimizes the difference in variances between groups; the partition represented by the square minimizes the difference in means between groups; the partition represented by the diamond maximizes $SSE_{kplus}$, and satisfying both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}

@@

The appendix includes additional figures that split the simulation results according to the other variables that varied between data sets (number of features; correlation between features; standard deviation of features; number of groups). Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased. Because the number of covariances increases quadratically when the number of features increases, it gets comparatively more difficult to fulfill the correlation objective with a higher number of features.

@@

This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using an interface function called `kplus_anticlustering()`, which is available from version 0.6.3 onward. The package can be installed in the R environment using the `install.packages()` command:

@@

This function call relied on the default k-plus setting that ensures similarity with regard to means and variances, but does not consider higher order moments or covariances.

@@

Table 4 illustrates the results that are obtained when using standard k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each. Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are still similar enough for any practical purpose---especially when considering inevitable measurement error in norming studies.

@@

Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity. In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity. K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and even covariances. Depending on the application and the needs of a user, a different subset of these objectives can be included in the anticlustering process. The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means. Interestingly, however, all k-plus variations basically do not change the original k-means criterion, because additional criteria can be incorporated by augmenting the input data.

>>>>

When optimizing a bicriterion objective such as $\mathit{SSE_{kplus}}$, the ideal outcome would be to find a partition where both the means and the variances are very similar between groups. However, research on multi-objective optimization has shown that cases are rare where a single solution satisfies all criteria to an optimal degree, making a tradeoff necessary [@ferligoj1992direct]. A motivating example illustrates that in the case of anticlustering, aiming to simultaneously optimize similarity of means and variances may however be feasible. I created 14 data points following a univariate normal distribution ($M = 15$, $SD = 3$) and then generated all `r n_partitions(14, 2)` possible ways to partition the 14 data points into $K = 2$ equal-sized groups. Figure 1 shows the scatterplot of the criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ across partitions. To divide the 14 elements into two groups that are similar with regard to their mean and variance, a partition should be chosen that is depicted in the upper right corner of the scatterplot. The partition that is represented by the squared symbol has the overall best highest $\mathit{SSE}$, but is outperformed by many other partitions with regard to $\mathit{SSE_{Var}}$. Similarly, the partition that maximizes $\mathit{SSE_{Var}}$ (the triangular symbol in Figure 1) is not well-suited to optimize $\mathit{SSE}$ at the same time. The diamond symbol represents the partition that maximizes $\mathit{SSE_{kplus}}$, i.e., the sum of $\mathit{SSE}$ and $\mathit{SSE_{Var}}$. This partition is located in the right upper corner and yields `r df$kmeans[which.max(df$kplus)] / max(df$kmeans) * 100`% of the global maximum $\mathit{SSE}$, and `r df$kvar[which.max(df$kplus)] / max(df$kvar) * 100`% of the global maximum value of $\mathit{SSE_{Var}}$. Thus, the optimal partition according to $\mathit{SSE_{kplus}}$ fulfills the two constituting criteria $\mathit{SSE}$ and $\mathit{SSE_{Var}}$ to an astonishing degree.

```{r, fig.cap = "The relationship between the objectives $SSE$ and $SSE_{Var}$ across all ways to partition $N = 14$ normally distributed random values into $K = 2$ groups. The partition represented by the triangular minimizes the difference in variances; the partition represented by the square minimizes the difference in means; the partition represented by the diamond maximizes $SSE_{kplus}$ and satisfys both criteria to an almost optimal degree.", fig.height = 5, fig.width = 5}

@@

The Appendix includes additional figures that split the simulation results according to the other variables that varied between simulation runs (number of features; correlation between features; standard deviation of features; number of groups). Here, the only meaningful moderating effect on anticlustering performance was that k-plus correlation was more strongly impaired than other methods when the number of features increased. Because the number of covariances increases quadratically when the number of features increases, it gets comparatively more difficult to fulfill the correlation objective with a higher number of features.

@@

This section demonstrates how readers can easily employ k-plus anticlustering using the R package `anticlust` [@papenberg2020] using the function `kplus_anticlustering()`, which is available from version 0.6.3 onward. The package can be installed in the R environment using the `install.packages()` command:

@@

@@

Table 4 illustrates the results that are obtained when using standard k-plus anticlustering to divide the stimulus set into 6 groups of size `r nrow(schaper2019) / 6` each. Even though the means and standard deviations are no longer perfectly matched between groups, arguably they are similar enough for any practical purposes.

@@

Anticlustering is a partitioning method that aims for high between-group similarity and within-group heterogeneity. In this paper, I presented the k-plus criterion for anticlustering, which focuses on maximizing between-group similarity. K-plus represents between-group similarity as discrepancies in distribution moments between groups: mean, variances, skewness, kurtosis, higher order moments, and covariances. Depending on the application, a different subset of these objectives can be included in the anticlustering process. The k-plus criterion is an extension of the classical k-means criterion, which only reflects how similar groups are with regard to their means. Interestingly, however, the k-plus extension does not change the original k-means criterion because additional criteria can be incorporated by augmenting the input data.
